---
title: "Lab 9 Intro"
date: "`r Sys.Date()`"
output: 
  ioslides_presentation: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(jpeg)
library(here)
library(janitor)
```

## Seating
1. If you already have a group, it will be helpful to start sitting next to each other.
2. If you don't have a group, hang tight we are working on it.
3. Please setup your computer as normal for today.

## Warm-up
1. In the data folder there is an epidemiology data set on an outbreak of malaria.
```{r}
malaria<-read_csv(here("lab9", "data","malaria.csv")) %>% clean_names()
malaria
```

2. `rdt` refers to rapid diagnostic test and they are identified here by age group.

3. Make the data tidy and store them as a new object.
```{r}
malaria_longer<-malaria %>% 
  pivot_longer(cols = starts_with("malaria_rdt"),
               names_to="age_class",
               values_to="cases") %>% 
  select(-malaria_tot) #select all columns except malaria total
malaria_longer
```



4. Which district had the highest *total* number of cases on July 30, 2020?
```{r}
malaria_longer %>% 
  filter(data_date=="2020-07-30") %>% 
  group_by(district) %>% 
  summarise(total_cases=sum(cases,na.rm=T)) %>% 
  arrange(-total_cases) %>% 
  head(n=1)

```


